---
title: "Clean Data"
output: html_notebook
---

# Optionally Install Packages, Load Packages

Below we will check whether you have the package called `pacman` installed on your computer. If not, we will install it. Then we will use that package to load the required packages to clean your data. These packages we will use are the tidyverse (for cleaning data),jsonlite (for unpacking lab.js data), janitor (for creating consistent column names), and here (for indexing your external data) packages.

```{r packages}
if(!require("pacman")) {install.packages("pacman")}
pacman::p_load("tidyverse", "jsonlite", "janitor", "here")
```

# Load Data

Ensure that your data is saved as a .csv file called "raw_data.csv" in the folder "data" at the root of this directory. If this is not the case, please rename your data file and put it in this folder. 

We will then load the data from this folder as follows:

```{r read-raw-data, message=FALSE}
raw_data <- read_csv(here("data", "raw_data.csv"))
```

# Clean Data

Next, we need to define where the .json data from lab.js lives in your data set. If you followed the step by step guide and renamed the embedded data to "labjs-data" in Qualtrics, this will work. 

If not, please ensure that you change the string in the chunk below to whatever you named your embedded data to. Details of this are in step 32 of the step by step guide.

```{r define-labjs-data-column}
labjs_column <- "labjs-data"
```

We will now clean up our data by taking the raw data, removing the extra rows Qualtrics puts into it, then replacing any missing data from lab.js with a format R expects to see. After this, we will export the data from the labjs-data column. Again, if this doesn't work for you have passed the correct name for the column with your lab.js data in it in the code chunk above. 

Next, we will tidy up our data by converting any data types to what R should expect (e.g. text = characters, numbers = numeric) and make sure the names are all consistent in their format.

```{r unpack-and-clean-raw-data, message = FALSE}
cleaned_data <- raw_data %>% 
  slice(3:n()) %>% # remove the first 2 rows which Qualtrics introduces
  # handle any missing data by making it an empty .json entry
  mutate(!!labjs_column := recode(.[[labjs_column]], .missing="[{}]")) %>%
  group_by_all() %>% # group the data by all variables
  do(fromJSON(.[[labjs_column]], flatten=TRUE)) %>% # extract .json data from the variables
  ungroup() %>% # remove grouping
  select(-matches(labjs_column)) %>% # remove column containing raw JSON
  type_convert() %>% # fix column data types
  clean_names() # fix names so they are consistent
```

# Repair Data

A few columns weren't calculated correctly or were stored in a messy format. We'll fix that here.

## Fix Consent

At this point, we may want to create a new column in our data called consent, which is made up of only the questions we have which correspond to giving consent in Qualtrics. The names for these columns are on the left for the block title containing your consent information.

For me, I had two consent questions, saved in Qualtrics as q2 and q3. I'll make a column called consent, which is only "Yes" if people give consent for both these questions. Otherwise, they get "No" because they haven't given full consent. We'll use this to filter only to people who gave consent later on.

If your Qualtrics columns have your consent questions under different titles, please change q2 and q3 here to be the titles you have. If you have additional consent questions, add in more & and the name of the column you have. For example `q2 == "Yes" & q3 == "Yes: & q4 == "Yes"`.

Finally, add any columns you used to make the consent column to the select call. This removes these columns from the data set.

```{r make-consent}
cleaned_data <- cleaned_data %>% 
  mutate(
    consent = case_when(
      q2 == "Yes" & q3 == "Yes" ~ "Yes",
      TRUE ~ "No"
    )
  ) %>% 
  select(-c(q2, q3))
```

# Save Data

Finally we will save our data back to the data folder, now under the name "cleaned_data.csv".

```{r save-cleaned-data}
write_csv(cleaned_data, here("data", "cleaned_data.csv"))
```
